---
title: "Strategy"
author: "Oskar Allerslev"
date: "2024-12-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Strategy

We assume that under the risk-neutral measure $\mathbb{Q}$, the underlying price $S_t$ follows;

$$
\begin{align*}
dS_t = S_t(r-q-\lambda \kappa)dt + S_t \sigma dW_t + S_{t-}(e^Y-1)dN_t
\end{align*}
$$
Where $r$ is the risk free rate. $q$ the continious dividend. $\lambda$ is the intensity of the jumps. $Y$ is the jump size. And lastly $\kappa = E^{\mathbb{Q}}(e^Y-1)=e^{\mu_J +\frac{1}{2}\sigma_J^2}-1$.

Under mertons model, we can write out the Europen Call option semi-analytically as:
$$
\begin{align*}
C_{Merton}(S_0,K,T)=\sum_{k=0}^\infty e^{-\lambda T} \frac{(\lambda T)^k}{k!}C_{BS}(S_0 e^{k\mu_J^*}, K, T, r,q, \sigma_{eff}^2)
\end{align*}
$$


```{r}

library(quantmod)
library(dplyr)
library(purrr)
library(fitdistrplus)
library(tibble)
source("optionchain.R")

```



# Example option chain

```{r}
options_df <- optionchain()
print(options_df)

```


```{r}
getSymbols("GERN", from="2019-01-01", to="2024-12-29")  
prices <- Cl(GERN)

#Compute daily log returns
returns <- diff(log(prices))[-1]  
print(returns)

```

# calibrate parameters for model 

```{r}
sd_ret <- sd(returns, na.rm = TRUE)
threshold <- 3 * sd_ret

jumps <- returns[abs(returns) > threshold]
no_jumps <- returns[abs(returns) <= threshold]

fit_jump <- fitdistrplus::fitdist(as.numeric(jumps), "norm")
mu_jump <- fit_jump$estimate[1]
sigma_jump <- fit_jump$estimate[2]

lambda <- length(jumps)/length(returns)

sigma <- sd(no_jumps, na.rm = TRUE)


print(c(mu_jump, sigma_jump, "sd no jump" = sigma, "lambda" =lambda))



```


```{r}
# Standard Black-Scholes (European) for a call:
BS_call_price <- function(S, K, T, r, q=0, sigma) {
  if(T <= 0) return(pmax(S - K, 0))
  d1 <- (log(S/K) + (r - q + 0.5*sigma^2)*T)/(sigma*sqrt(T))
  d2 <- d1 - sigma*sqrt(T)
  call <- S*exp(-q*T)*pnorm(d1) - K*exp(-r*T)*pnorm(d2)
  return(call)
}

# Merton's jump-diffusion call
merton_call_price <- function(
  S0, K, T, r, q, sigma,  # continuous vol
  lambda, muJ, sigmaJ,    # jump intensity & jump dist parameters
  max_terms=50
) {
  # Summation over k jumps
  sumval <- 0
  # e^-lambdaT factor
  e_negLT <- exp(-lambda * T)
  
  for(k in 0:max_terms) {
    # Poisson weight
    w_k <- e_negLT * ((lambda*T)^k / factorial(k))
    
    # Shifted underlying for k jumps:
    # Often: S_k = S0 * exp(k * muJ)
    # Some references include correction factors (e.g. -0.5*sigmaJ^2*k),
    # but let's keep the simpler standard form:
    S_k <- S0 * exp(k * muJ)
    
    # Price a vanilla call using BS with the same sigma. 
    # (In some derivations you modify sigma => but let's keep it simple.)
    call_k <- BS_call_price(S_k, K, T, r, q, sigma)
    
    sumval <- sumval + w_k * call_k
  }
  
  return(sumval)
}

```


```{r}

options_df <- optionchain()


na.omit(optionchain())

filtered <- options_df %>%
  dplyr::filter(
    expiration_date == as.Date("2025-01-17"),
    bid != 0,
    ask != 0
  )
# View the result
print(filtered)

```


```{r}
today <- as.Date("2024-12-29")
expiry <- as.Date("2025-02-21")
T_yr <- as.numeric(expiry - today)/365
S0_gern <- as.numeric(last(prices))
r <- 0.04
q <- 0.00




stock <- filtered %>%
  mutate(
    mid_price = (ask + bid)/2,
    merton_theo = map_dbl(
      strike,
      ~ merton_call_price(
        S0_gern, .x, T_yr, r, q,
        sigma,
        lambda,
        mu_jump,
        sigma_jump
      )
    ),
    mispricing = mid_price - merton_theo
  )



mispriced <- data.frame(
  ticker = stock$ticker,
  expiration = stock$expiration_date,
  type = stock$type,
  strike = stock$strike,
  ask = stock$ask,
  bid = stock$bid,
  mid_price = stock$mid_price,
  model_theoretical = stock$merton_theo,
  mispricing = stock$mispricing,
  spread = stock$ask - stock$bid
  
)

print(mispriced)


```



- **4.1 If You’re Bullish on GERN**

Buy calls with negative mispricing (underpriced). 
Or,Sell puts with positive mispricing (overpriced).

But watch the potential for large margin requirements if you’re short puts without a hedge.

- **4.2 If You’re Bearish on GERN**

Buy puts with negative mispricing (cheap).
Sell calls with positive mispricing (expensive).

Again, short calls can have unlimited upside risk unless you do a covered call (own the underlying) or a call spread.
































